---
title: "EDS241: Assignment 04 - Price Elasticity"
author: "Julia Parish"
date: "2022/03/11"
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, echo = TRUE, message = FALSE, warning = FALSE)

options(scipen = 999) # not scientific notation

```

```{r}
# load packages
packages = c("stargazer", 
             "here",
             "tidyverse",
             "stringr", 
             "janitor", 
             "cowplot", 
             "tinytex", 
             "datasets", 
             "tibble",
             "rio",
             "car",
             "readxl",
             "estimatr",
             "patchwork", 
             "kableExtra",
             "huxtable")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

```

# EDS241 Environmental Policy Evaluation Assignment 04

\noindent This statistical analysis was completed as an assignment for the course, Environmental Data Science 241: Environmental Policy Evaluation. 

This question will ask you to estimate the price elasticity of demand for fresh sardines across 56 ports located in 4 European countries with monthly data from 2013 to 2019. 

# Homework Questions

## Question A:

\noindent Estimate a bivariate regression of log(volume_sold_kg) on log(price euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.


### Answers\

\noindent


## Question B:

\noindent Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic.


### Answers\

\noindent

## Question C:

\noindent Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines?


### Answers\

\noindent


## Question D:

\noindent Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint: you can use the command “as.factor(country) + as.factor(year) +as.factor(month)” to the ivreg function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.


### Answers\

\noindent