---
title: "EDS241: Assignment 02 - Air Quality Regulations"
author: "Julia Parish"
date: "2022/02/04"
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, echo = TRUE, message = FALSE, warning = FALSE)

options(scipen = 10) # not scientific notation

```

```{r, include = FALSE}
# load packages
packages = c("stargazer", 
             "here", 
             "tidyr", 
             "dplyr",
             "stringr", 
             "janitor", 
             "cowplot", 
             "ggplot2", 
             "tinytex", 
             "datasets", 
             "tibble",
             "rio",
             "car")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

```

# EDS241 Environmental Policy Evaluation Assignment 02

\noindent This statistical test was completed as an assignment for the course, Environmental Data Science 241: Environmental Policy Evaluation. The goal of this assignment was to provide a simple test of whether the effects of air quality regulations are the same across various regions with diverse demographics. the test will analyze the [NOx Budget Program](https://www.epa.gov/airmarkets/nox-budget-trading-program), a cap-and-trade market for nitrogen oxides (NOx) emissions from power plants lead to similar effects in counties that are predominantly white versus counties that are predominantly African American. 

**Source:** 
The data was sourced from researched conducted by [Dr. Deschenes et al.](https://olivierdeschenes.weebly.com/uploads/1/3/5/0/135068654/defensive-investments-and-the-demans-for-air-quality.pdf).

**Glossary**
**FIPS** - Federal Information Processing System codes identifying each county
**NBP** - indicator =1 if the county was regulated under the NOx Budget Program)
**PctBlack**(fraction of the county population that is African American)
**Dnox_masstons** (change in annual NOx emissions from all power plants in a county between 2000 and 2008 (in tons)).

## Question A. Make a histogram depicting the distribution of Dnox_masstons.


## Question B. Create an indicator =1 if the county has PctBlack above the sample median, and =0 otherwise (in the rest of the assignment, I refer to this variable as ‘D’). What is the average of PctBlack for counties above the median, i.e. counties for which D=1?


## Question C. Estimate a regression of Dnox_masstons on NBP. Interpret the estimated intercept and the coefficient on NBP.


## Question D. Create an interaction between the variables NBP and D. Estimate a regression of Dnox_masstons on NBP, D, and this interaction. Interpret each estimated regression coefficient, including the intercept.


## Question E. What is the predicted Dnox_masstons in a county that was not regulated under NBP and where PctBlack is above the sample median (i.e., where D=1)? Report the 95% confidence interval for this prediction. Make sure to use “heteroskedasticity-robust” standard errors.




