---
title: "EDS241: Assignment 03 - National Natality Detail"
author: "Julia Parish"
date: "2022/02/20"
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
---

```{r setup, include=FALSE}
# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, echo = TRUE, message = FALSE, warning = FALSE)

options(scipen = 999) # not scientific notation

```

```{r, include = FALSE}
# load packages
packages = c("stargazer", 
             "here", 
             "tidyr",
             "tidyverse",
             "dplyr",
             "stringr", 
             "janitor", 
             "cowplot", 
             "ggplot2", 
             "tinytex", 
             "datasets", 
             "tibble",
             "rio",
             "car",
             "readxl",
             "estimatr",
             "patchwork", 
             "kableExtra",
             "huxtable")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

```

# EDS241 Environmental Policy Evaluation Assignment 03

\noindent This statistical analysis was completed as an assignment for the course, Environmental Data Science 241: Environmental Policy Evaluation. It is an application of estimators based on treatment ignorability. The goal of this assignment was to to estimate the causal effect of maternal smoking during pregnancy on infant birth weight using the treatment ignorability assumptions. The data are taken from the National Natality Detail Files and the data files for this assignment is a random sample of all births in Pennsylvania during 1989-1991. Each observation is a mother-infant pair.

\noindent The outcome and treatment variables are:

- `birthwgt` = birth weight of infant in grams 
- `tobacco` = indicator for maternal smoking

\noindent The control variables are:

- `mage`: mother's age
- `meduc`: mother's education
- `mblack`: = 1 if mother is Black
- `alcohol`: = 1 if consumed alcohol during pregnancy
- `first`: = 1 if first child
- `diabete`: = 1 if mother is diabetic
- `anemia`: = 1 if mother anemic 

\noindent *Note: This exercise asks you to implement some of the techniques presented in Lectures 6-7. This homework is a simple examination of these data. More research would be needed to obtain a more definitive assessment of the causal effect of smoking on infant health outcomes. Further, for this homework, you can ignore the adjustments to the standard errors that are necessary to reflect the fact that the propensity score is estimated. Just use heteroskedasticity robust standard errors in R. If you are interested, you can read Imbens and Wooldridge (2009) and Imbens (2014) for discussions of various approaches and issues with standard error estimations in models based on the propensity score.*

# Data
```{r, message=FALSE}
# read in the data
nn_data <- read_csv(here("hw03/data/smoking.csv")) %>% 
  clean_names()
```

\newpage

# Homework Questions

## Question A:

\noindent What is the unadjusted mean difference in birth weight of infants with smoking and non-smoking mothers? Under what assumption does this correspond to the average treatment effect of maternal smoking during pregnancy on infant birth weight? Provide some simple empirical evidence for or against this assumption.
For the last part of (a), regress your favorite covariate on the smoking status of mothers. For example, think of regressing meduc ~ tobacco. Is the mean difference in the education level of smoking and non-smoking mothers statistically different from zero? What does that say about the required assumption to interpret the unadjusted mean difference as causal? 

```{r , include=TRUE}
# calculate the unadjusted mean difference in birth weight of infants with smoking and non-smoking mothers

smoker <- nn_data %>% filter(tobacco == 1)
nonsmoker <- nn_data %>% filter(tobacco == 0)

smoker_mean <- round(mean(smoker$birthwgt), 3)
nonsmoker_mean <- round(mean(nonsmoker$birthwgt), 3)

unadj_diff <- smoker_mean - nonsmoker_mean
```

```{r}
# Regress infant birth weight (birthwgt) in grams on the indicator for maternal smoking (tobacco)
mod_a1 <- lm_robust(birthwgt ~ tobacco, data = nn_data)
```

```{r}
#create table with regression results
mod_a1_table <- tidy(mod_a1)

mod_a1_table %>% 
  select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  kable()

mod_a1_table
```

```{r}
# Regress the education level (meduc) of the mother on the indicator for maternal smoking (tobacco)
mod_a2 <- lm_robust(meduc ~ tobacco, data = nn_data)
```

```{r}
#create table with regression results
mod_a2_table <- tidy(mod_a2)

mod_a2_table %>% 
  select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  kable()

mod_a2_table
```

### Answers\

\noindent The mean infant birthweight for those born from mothers who do smoke is `r smoker_mean`. The mean infant birthweight for those born from mothers who do NOT smoke is `r nonsmoker_mean`. The unadjusted mean difference in birth weight of infants with smoking and non-smoking mothers is `r unadj_diff` grams. This means that, on average, infants born to mothers who smoke weighed `r abs(round(mod_a1[[1]][[2]], 2))` grams less than infants born to mothers who do not smoke. 

\noindent The unadjusted mean difference corresponds to the average treatment effect of mother's smoking during pregnancy on infant birth weight, assuming that the treatment of whether a mother is a smoker or not is randomly assigned and statistically significant. Smoking status of mothers during pregnancy is independent of `Y(1)` and `Y(0)`.

\noindent Empirical evidence against the assumption that smoking treatment is randomly assigned to mothers during pregnancy is that another variable, education, is significantly correlated with the indicator for maternal smoking as shown in the linear regression model, `mod_a2`.

## Question B:

\noindent Assume that maternal smoking is randomly assigned conditional on the observable covariates listed above. Estimate the effect of maternal smoking on birth weight using a linear regression. Report the estimated coefficient on tobacco and its standard error.

```{r}
# Regress infant birth weight (birthwgt) in grams conditional on all variables in the data set

mod_b <- lm_robust(birthwgt ~ ., data = nn_data)
```

```{r}
#create table with regression results
mod_b_table <- tidy(mod_b)

mod_b_table %>% 
  select(term, estimate, std.error, p.value, conf.low, conf.high) %>%
  kable()

mod_b_table
```

### Answers\

The estimated effect of maternal smoking on birth weight is a decrease of `r abs(round(mod_b[[1]][[4]], 2))` grams on average. The standard error is `r abs(round(mod_b[[2]][[4]], 2))`. 

## Question C:

\noindent Use the exact matching estimator to estimate the effect of maternal smoking on birth weight. For simplicity, consider the following covariates in your matching estimator: create a 0-1 indicator for mother's age (=1 if mage>=34), and a 0-1 indicator for mother's education (1 if meduc>=16), mother's race (mblack), and alcohol consumption indicator (alcohol). These 4 covariates will create 2 * 2 * 2 * 2 = 16 cells. Report the estimated average treatment effect of smoking on birthweight using the exact matching estimator and its linear regression analogue (Lecture 6, slides 12-14).
Once you have your 4 dummy variables, you can create a group variable `g` using the paste0() function. For example, `mutate(g = paste0(d1,d2,d3,d4))` . The resulting `g` will include all potential observed combinations of the 4 dummy variables in the data. You can then control for `factor(g)` in the regression model. 
To calculate the exact matching estimator, use this `g` grouping variable and the code from lines 76 to 97 in the `TIA.R` script on gauchospace. In this case, Y = birthwgt, X = g, D = tobacco. Since we observe `Y1` or `Y0`, you can ignore line 77. *see TIA Table #ydiff = delta for x, w_ATE # of obs for rows, w_ATT = weights* 

### Section 1: Exact matching estimator

\noindent Use the exact matching estimator to estimate the effect of maternal smoking on birth weight. Consider the following covariates in your matching estimator: 

- mother's age (=1 if mage>=34), 
- mother's education (1 if meduc>=16), 
- mother's race (mblack), and 
- alcohol consumption indicator (alcohol). 

These 4 covariates will create 2 * 2 * 2 * 2 = 16 cells.

```{r}
# create 0-1 indicators for mother's education and age.

matching_nn_data <- nn_data %>% 
  mutate(
    mage_sq = (mage*mage),
    mage = case_when(
      mage >= 34 ~ 1,
      mage <34 ~ 0),
    meduc = case_when(
      meduc >= 16 ~ 1, 
      meduc < 16 ~ 0), 
    mblack = as.factor(mblack), 
    alcohol = as.factor(alcohol),
    covariates = paste0(mage, meduc, mblack, alcohol)
    )
```

```{r}
# create average treatment estimate of smoking on birth weight using exact matching estimator

tia_table <- matching_nn_data %>%
  group_by(covariates, tobacco) %>% 
  summarise(n_obs = n(), # number of observations
            birthwgt_mean = mean(birthwgt, na.rm = TRUE)) %>% # calculate birthwgt mean by X by treatment cells
  gather(variables, values, n_obs:birthwgt_mean) %>% # reshape the dataframe
 
  mutate(variables = paste0(variables, "_", tobacco, sep = "")) %>%  # combine the treatment and variables for reshaping
  pivot_wider(id_cols = covariates,  # reshape data by treatment and X cell
              names_from = variables, 
              values_from = values) %>% 
  ungroup() %>%  
  mutate(birthwgt_diff = birthwgt_mean_1 - birthwgt_mean_0,  # calculate birthwgt_diff
         w_ATE = (n_obs_0 + n_obs_1) / (sum(n_obs_0) + sum(n_obs_1)), # calculate ATE
         w_ATT = n_obs_1 / sum(n_obs_1)) %>%  # calculate ATT weights
  mutate_if(is.numeric, round, 2)

```

```{r}
stargazer(tia_table, type= "text", summary = FALSE, digits = 2)
```

```{r}
# MULTIVARIATE MATCHING ESTIMATES OF ATE AND ATT
ate = sum((tia_table$w_ATE)*(tia_table$birthwgt_diff))
ate 

att = sum((tia_table$w_ATT)*(tia_table$birthwgt_diff))
att
```


#### Answers\

The average treatment effect of smoking on birthweight using the exact matching estimator is `r ate` grams.




## Question D:

\noindent Estimate the propensity score for maternal smoking using a logit estimator and based on the following specification: mother’s age, mother’s age squared, mother’s education, and indicators for mother’s race, and alcohol consumption. glm(formula, family = binomial(), data) is a logit model.




### Answers\



## Question E:

\noindent Use the propensity score weighted regression (WLS) to estimate the effect of maternal smoking on birth weight (Lecture 7, slide 12). # see cgl.R, glm is a logit







### Answers\




